---
title: "Project 3: STAT302PACKAGE Tutorial"
author: "Rolina Cao"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302PACKAGE Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(STAT302PACKAGE)
```

# Introduction

This package contains four functions:\
* my_t_test : run a one sample t-test\
* my_lm : fit data into a linear model\
* my_knn_cv : take n-nearest neighbors as model to run k-fold cross validation.\
* my_rf_cv : perform random forest cross validation\

<br>

To install and use this package, run the following lines of code in console:

```{r, eval=FALSE}
# install and load package
devtools::install_github("RolinaC/STAT302PACKAGE")
library(STAT302PACKAGE)
```

Here are what we need for data
```{r}
data("my_penguins")
data("my_gapminder")
```

# my_t.test Tutorial

Here we make a reference to the 'lifeExp' data from 'my_gapminder' to show how the my_t.test functions. Note that the 'lifeExp' indicates the life expectancy at birth.

### Two-sided Test

First, we demonstrate a test of the hypothesis that is two-sided and the null
hypothesis is 60 indicating the mean life expectancy is 60 years; meanwhile, the alternative hypothesis holds that the mean life expectancy is not 60 years.

```{r}
two_side <- my_t_test(my_gapminder$lifeExp, "two.sided", 60)
```
The p-value, `r two_side$p_val`, is 


Second, we demonstrate two single-tailed tests of the hypothesis, one lower tail and one upper tail. Both has the null hypothesis saying 60 indicating the mean life expectancy is 60 years anf the alternative hypothesis holds that the mean life expectancy is not 60 years.

```{r}
low <- my_t_test(my_gapminder$lifeExp, "less", 60)
up <- my_t_test(my_gapminder$lifeExp, "greater", 60)
```
